Biases in Planet Occurrence Caused by Unresolved Binaries in Transit Surveys

### 2018-04-13 paper/review/response_1.txt

Dear Dr. Sigurdsson,

We thank you for organizing the review of our paper, and the referee
for constructive criticism.  We have considered the referee's remarks
carefully and revised our manuscript accordingly.

Below, we have reproduced the referee's report, along with our
point-by-point responses. A list of changes is also appended to the
manuscript.

Sincerely,

Luke Bouma, Kento Masuda, and Joshua Winn.

********************************************************************************

Referee comment:

Section 2, end of paragraph 2, For each of the 4 points regarding missed
transit signals, how often do they occur? (could add references).

  Response:

  We cannot identify "4 points regarding missed transit signals" in section 2,
  end of paragraph 2.

  Assuming that the referee is referring to points #1, #2, #3, enumerated at
  the end of section 2, paragraph 4:

  We do not know how often each of these effects happen -- the goal of this
  section is simply to introduce the existence of these sources of uncertainty.
  Could the referee please clarify their request, or perhaps give some indication
  of which references we might add?

********************************************************************************

Referee comment:

Section 3.1, first sentence, "Since the effects of binarity are most pronounced
when the two stellar components are similar" <- please also point to the place
in the text where this is explained.

  Response:

  The original idea was that binarity produces the largest errors in inferred
  planetary radii when the two stellar components are similar.  However, this
  is only true if the detected planet orbits the primary.  If it orbits the
  secondary, it is false.

  We have accordingly reworded this sentence.

********************************************************************************

Referee comment:

Section 3.2, following equations 4, you may want to change "stars" to "host
star".

  Response:

  Equation 4 is in Section 3.1, and the word "stars" does not appear around
  that equation in an immediately obvious way.

  Perhaps the referee meant around equation 14 of section 3.2?

  We have not taken any action on this point.

********************************************************************************

Referee comment:

Section 4.1, when exploring the Howard et al. planet radius distribution, how
sensitive are the results to the choice of gamma? Do mathematical artifacts
here make the results insensitive to the choice of the value of gamma?

  Response:

  Consider the cases of γ=-2.5, and γ=-3.5. (Well outside the power laws needed
  to fit Howard et al's results).  Applying Equations 28 and 30, we find:

                   γ=-2.5      γ=-2.92      γ=-3.5
  --------------------------------------------------
  Equation 28
  Γ_a/Γ_0           1.023        1.003      0.980

  Equation 30 
  Γ_a/Γ_0           1.095        1.048      0.993

  We have added a sentence summarizing this result.

********************************************************************************

Referee comment:

Section 4.4, end of paragraph 1, Add Lopez and Fortney citation.

  Response:

  We have corrected this omission.  In addition to referencing the work by
  Lopez & Fortney 2013, we now also refer to that by Owen & Wu 2013.

********************************************************************************

Referee comment:

Section 4.4, paragraph 2, "radius distribution is assumed to have a complete
absence of planets with sizes between 1.5 and 2R_Earth", might clarify that the
"true" radius distribution is assumed to have this feature after fitting the
apparent distribution.

  Response:

  We have clarified the wording in this paragraph, in a manner that should
  clarify the distinction between Fulton+17's apparent distribution, and the true
  distribution that we assume for this numerical experiment.

********************************************************************************

Referee comment:

Section 4.4, paragraph 2, More discussion of the spurious feature below 1.5
R_Earth would be helpful.

  Response:

  The peak in apparent occurrence rate density between 1.3 and 1.5 R_earth
  (solid blue line relative to dotted light blue line) is the contribution from
  secondaries with true radii greater than 2R_earth.

  We have added a comment to this effect in the caption of Figure 6. We have
  also increased the resolution of the grid in planet radius, to emphasize that
  this is a feature, rather than a bug.

  [
  OPTIONAL:
  As an aside, consider the numerical experiment of a single planet size, and a
  distribution of binary light ratios. I.e., imagine the true occurrence rate
  density is a delta function around r_p. The resulting apparent occurrence rate
  density would have contributions from primaries between r_a = r_p/sqrt(2) and
  r_a = r_p.  It would have contributions from secondaries from (roughly) r_a =
  0 up to r_p/sqrt(2). We ran this numerical experiment, and though did not
  mention it in the manuscript, we found that the contribution from secondaries
  peaked at r_a = r_p/sqrt(2).

  This delta function peak is analogous to what is seen in Figure 6.
  ]

********************************************************************************

Referee comment:

Section 4.4, paragraph 2, "may be even more devoid of planets than it appears",
might want to emphasize again here that this conclusion is very dependent on
the true distribution model plugged into the formalism.

  Response:

  We agree that the apparent rate density distribution is very dependent on
  the assumed true rate density distribution. However, we have opted not to
  change our original wording, which we believe accurately states our
  conclusion.

  Our conclusion here is that for an assumed true rate density distribution
  that roughly matches observations, there is no way for binarity to
  cause underestimates of the rate density inside a "dip" of planet occurrence.

  Restated, if there is a "gap" in the true rate density distribution,
  the apparent rate density distribution will always fill it in.

  Thus, if a gap is observed in an apparent distribution, the true distribution
  must have a gap at least as deep, if not deeper.

********************************************************************************

Referee comment:

Section 5, Hot Jupiter paragraph, states that result of this study find that
unresolved binaries likely result in an overestimated of hot-Jupiter occurrence
rates, implied by the Fulton gap model in Section 4.4. Previously, in section
4.2, paragraph 3, it states the occurrence rate density of planets greater than
8 Earth radii are underestimated. The abstract just says net systematic error
of order 5%. It would help to include a table describing the different tests in
this study, and how parameters are varied (how knobs are turned).

  Response:

  In Section 5, Hot Jupiter paragraph, we erroneously referred to Section 4.4;
  this should have been a reference to Section 4.5 (the gaussian radius
  distribution model).  We have corrected the error.

  We have now also added a table summarizing the case studies, as well as
  the parameters that are varied.

********************************************************************************

Referee comment:

- Recommend using the term "uncertainty" instead of "errors" .

  Response:

  We have changed our terminology accordingly in many places throughout the
  manuscript (see list of changes).

********************************************************************************

Referee comment:

- May want to comment on parameterizing the occurrence rate in terms of planet
  radius and insolation flux. The semi major axis will depend on which star in
  the binary the planet orbits.

- May want to comment on paper by Danley Hsu: Occurrence rates from Kepler
  using approximate Bayesian computation. You could speculate on what their
  data generating process would yield when plugged into this formalism.

- May want to comment on the parameter space and telescopes for resolving
  binaries of various separations, etc. and the probability of being able to
  resolve binaries with followup.

  Response:

  We have added a "Caveats and Outlook" paragraph that touches on all three of
  these issues.

### 2018-04-13 paper/main.tex
aj submitted
add response to referee comments

\documentclass[12pt,modern]{aastex61}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{grffile} % checks for multiple dots in figure filenames
\usepackage{empheq} % for boxing equations
\newcommand*\widefbox[1]{\fbox{\hspace{0.2cm}#1\hspace{0.2cm}}}

@@

\title{ Biases in Planet Occurrence Caused by Unresolved Binaries in
        Transit Surveys }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\affiliation{
    NASA Sagan Fellow}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic errors in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the errors are potentially larger: the occurrence of Earth-sized
planets could be overestimated by as much as 50\%. We also show that
whenever high-resolution imaging reveals a transit host star to be a
binary, the planet is usually more likely to orbit the primary star
than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

A lingering concern about these studies is that in most cases,
investigators have assumed that all of the targets in the survey
are single stars~\citep[\textit{e.g.},][]{
  howard_planet_2012,fressin_false_2013,
  dressing_occurrence_2015,burke_terrestrial_2015}. 
In reality, many of the sources that are monitored in a transit survey are
unresolved multiple-star systems, mainly binaries.  Unrecognized
binaries cause numerous systematic errors in the planetary occurrence
rates.  For example, when there is a transiting planet around a star
in a binary, the additive constant light from the second star reduces
the fractional loss of light due to the planet.  This makes transit
signals harder to detect and lowers the number of detections.  On the
other hand, a binary system presents two opportunities to detect
transiting planets, which could increase the overall number of
detections.

@@

There are many potential pitfalls in this calculation.  Some genuine
transit signals are missed even if they formally exceed the
signal-to-noise threshold, because of the probabilistic nature of
transit detection.  Planets can be misclassified due to statistical
and systematic errors in the catalogued properties of the stars.  Some
transit-like signals are spurious, arising from noise fluctuations or
failures of ``detrending'' the astrophysical or instrumental
variations in the photometric signal.  Poor angular resolution leads
to blends between eclipsing binary stars and other stars along nearly
the same line of sight, producing signals that mimic those of
transiting planets.

Here, though, we will focus exclusively on problems that arise from
the fact that many stars exist in unresolved binary systems.  We will
also focus on the errors in planet occurrence as a function of radius,
rather than orbital period. This is because when more than one transit
is detected (as is usually required by the surveyors), the orbital
periods can be measured without ambiguity regardless of whether the
host star is single or one member of a binary.  Even with this narrow
focus, there are numerous sources of error.  All three of the
quantities in Equation~\ref{eq:occ_rate_simple} are biased:
\begin{enumerate}
%    
    \item The number of detected planets, $N_{\det}$, is actually the
      number of detected planets that {\it appear} to have size $r$,
      orbiting stars that {\it appear} to have mass $M$.  Whenever the
      planet-hosting star is part of a binary,
%    
    \begin{itemize}
        \item the planet's size could be misclassified because of the
          reduction in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified
          because its light is combined with a second star of a
          different spectral type.
%        
    \end{itemize}
%    
    \item The number of searchable stars, $N_\star$, is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of
          the secondary stars that were inadvertently searched for
          transiting planets;
%        
        \item toward higher values, because some of the stars that
          appeared to be searchable are in fact binaries for which the
          amplitude of the photometric signal would have been reduced
          to an undetectable level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$ is biased because the
      planet-hosting star could be misclassified.  At fixed orbital
      period, the transit probability is proportional to
      $\rho^{-1/3}$, where $\rho$ is the stellar mean
      density~\citep{Winn2010}.  Therefore, any errors in determining
      the host star's mean density lead to errors in the correction
      for the transit probability.
%    
\end{enumerate}

@@

Since the effects of binarity are most pronounced when the two stellar
components are similar, we begin by considering a universe in which
all stars are identical, with mass $M$, radius $R$, and luminosity
$L$.  Single stars are uniformly distributed in space with a number
density of $n\s$ stars per cubic parsec, and binary systems are
uniformly distributed with number density $n\b$.  In this scenario,
stars are never misclassified because the combined light of a binary
has the same color and spectrum as a single star.  We further assume
that all planets have the same radius, $r$, and occur around single
stars and members of binaries at the same rate, $\Lambda_r$.

@@

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=0.6\textwidth]{figures/f1.pdf}
    \end{center}
    \caption{ {\it Top.}---Volumes within which sources appear to be
    searchable for planets of radius $r$.  Single stars are searchable
    out to a distance $d_0$, at which point they become too faint to
    allow the detection of transits.  Binaries are brighter and
    therefore appear to be searchable out to a larger distance.  {\it
    Bottom.}---Volumes within which stars appear to be searchable for
    planets of radius $r/\sqrt{2}$.  }
    \label{fig:model_1_volumes}
\end{figure}

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/f2.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{ Apparent occurrence rate $\Lambda\a$, and true
    occurrence rate $\Lambda$ in a universe with only one type of
    star, one type of planet (with radius $r$), and $n\b/n\s=0.05$,
    for $n\b$ the number density of binary systems, and $n\s$ the
    number density of single stars.  The occurrence rate of planets
    with radius $r$ is underestimated, and the occurrence rate of
    planets with radius $r/\sqrt{2}$ is overestimated.}
    \label{fig:occ_rate_model_1}
\end{figure}

@@

We can now gauge the size of the systematic errors associated
with unresolved binaries, for cases of interest.  Throughout this
section we assume $R\propto M$, and $L \propto M^{\alpha}$ where
$\alpha = 3.5$, as is roughly the case for main-sequence stars.
With these assumptions, Equations~\ref{eq:dil_1} and~\ref{eq:dil_2}
become
\begin{equation}
  \mathcal{D}_1
  =  (1+q^\alpha)^{1/2}~~{\rm and}~~\mathcal{D}_2
  = q(1+q^{-\alpha})^{1/2}.
\end{equation}

@@

\begin{figure}[!tb]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/f3.pdf}
    \caption{ The mass ratio distribution of binaries in a
    magnitude-limited sample, assuming the underlying volume-limited
    distribution is uniform and the luminosity-mass relation is
    $L\propto M^{3.5}$. For transit signals of a given amplitude, the
    sample of searchable sources is magnitude-limited, causing high
    mass ratios to be over-represented. }
    \label{fig:q_distribn_mag_limited}
\end{figure}

@@

Again assuming $\Gamma_0=\Gamma_1=\Gamma_2$, the apparent occurrence
rate density is 
\begin{align}
    %see annoying_q_factor_dec22.pdf for derivation
    \notag
    %\frac{\left.\Gamma\a(r\a)\right|_{r\a= r}}{\Gamma_0(r)} 
    \frac{\Gamma\a(r\a)}{\Gamma_0(r\a)} 
    &=
    \frac{1}{1+\mu}
    \left[
      1 + \frac{1}{\mathcal{N}_q} \frac{n\b}{n\s}
    \left(
      \int_0^1 {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\gamma+4}{2}} +
      \right.
      \right. \\
      %isn't latex great?
      &\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
      \left.\left.
      \int_0^1 {\rm d}q\,q^{\beta+\gamma+\frac{5}{3}} 
      (1+q^\alpha)^{\frac{3}{2}}
      (1+q^{-\alpha})^{\frac{\gamma+1}{2}}
    \right)
    \right],
    \label{eq:powerlaw_vary_binary}
\end{align}
Adopting the realistic numerical values $n\b/n\s=0.79$, $\alpha =
3.5$, $\beta=0$, and $\gamma=-2.92$, the summed integrals in
Equation~\ref{eq:powerlaw_vary_binary} give $(\ldots)\approx 1.503$,
and $\Gamma\a/\Gamma_0 = 1.048$.  The apparent occurrence rate density
is 4.8\% higher than the true value.  We conclude that for planets around
Sun-like stars with orbits within 0.25\,AU and radii in the range
2--17\,$R_\oplus$, the systematic errors associated with binarity are
on the order of a few percent.

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/f4a.pdf}{figures/f4b.pdf}
    \caption{ {\it Left.}---Apparent occurrence rate density
      ($\Gamma\a$), compared to the true occurrence rate density
      ($\Gamma_0$) for single stars and primary stars in binaries.
      Three different cases are plotted, with different choices for
      the number of planets per secondary star relative to
      that for single stars ($Z_2/Z_0$).
      {\it Right.}---Same, integrated over $0.5\,R_\oplus$
      bins to give apparent and true occurrence rates.
      In both cases the true planet
      radius distribution is specified by
      Equation~\ref{eq:model3_radius_distribution}.
    }
    \label{fig:occ_rate_model_3_log}
\end{figure}

For planets smaller than 2\,$R_\oplus$, the true occurrence rate
density is more uncertain because such planets are more difficult to
detect.  To investigate the systematic errors associated with binarity
on the inferred occurrence rate of Earth-sized planets, we need to
make plausible assumptions about the true occurrence rate density.
Obviously if the rate density continues to vary as $r^{-3}$ down to
much smaller planet sizes, the results of the preceding section will
hold.  But there is no particular reason to think this will be the
case and indeed, some investigators have concluded that the rate
density begins to level off to a constant value as $r$ decreases below
$2R_\oplus$~\citep{petigura_plateau_2013}.  To investigate the
implications we consider a broken power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2R_\oplus \\
        {\rm constant} & \text{for } r\leq2R_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
In this case the integrals that appear in
Equation~\ref{eq:general_Gamma_a} are tedious to work out
analytically, leading us to evaluate them numerically.\footnote{We
refer the interested reader to our online code for performing this
computation: \url{github.com/lgbouma/binary_biases}, commit
\texttt{fd41b4b}.} Figure~\ref{fig:occ_rate_model_3_log} shows the
results, again for the case $n\b/n\s=0.79$, $\alpha = 3.5$, $\beta=0$,
and $\gamma=-2.92$.

@@

\subsection{Which star has the planet?}
\label{subsec:whichstar}
\begin{figure}[!t]
    \centering
    \includegraphics[width=\textwidth]{figures/f5.pdf}
    \caption{ Fraction of detected planets that orbit single stars
      (solid lines), primary stars of binaries (dashed lines), and
      secondary stars of binaries (dotted lines), as a function of the
      apparent radius of the planet.  We assume that single stars and
      primaries have the same number of planets per star.  Three
      different cases are plotted, differing in the relative
      occurrence of planets around secondary stars ($Z_2/Z_0$).  The
      true radius distribution is given by
      Equation~\ref{eq:model3_radius_distribution}.}
    \label{fig:frac_model_3}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/f6a.pdf}{figures/f6b.pdf}
    \caption{ {\it Left.}---True and apparent occurrence rate densities,
      and {\it right.}---occurrence rates, as a function of planet
      radius, using a model for the radius distribution that exhibits
      a gap between 1.5 and 2\,$R_\oplus$
      (Equation~\ref{eq:model4_radius_distribution}). }
    \label{fig:model_4}
\end{figure}

\citet{fulton_california-_2017} recently reported a ``gap'' in the
radius distribution of close-in planets around Sun-like stars, between
planet sizes of 1.5 and 2\,$R_\oplus$.  This can also be visualized as
a ``valley'' in the occurrence rate density as a function of radius
and period.  The existence of the gap has been independently
corroborated from a sample of KOIs with asteroseismically-determined
stellar parameters~\citep{van_eylen_asteroseismic_2017}.  Such a
feature had been predicted as a consequence of the gradual
photo-evaporation of the hydrogen-helium atmospheres of small rocky
planets, during the first 100~Myr of the age of the system when the
host star produces a higher flux of ultraviolet and X-ray
radiation~\citep{owen_evaporation_2017}.

>>>>

\documentclass[12pt,modern,trackchanges]{aastex61}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{grffile} % checks for multiple dots in figure filenames
\usepackage{empheq} % for boxing equations
\newcommand*\widefbox[1]{\fbox{\hspace{0.2cm}#1\hspace{0.2cm}}}

@@

\title{ Biases in Planet Occurrence Caused by Unresolved Binaries in
        Transit Surveys }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\affiliation{
    NASA Sagan Fellow}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic 
\deleted{errors}\added{uncertainty} in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the errors are potentially larger: the occurrence of Earth-sized
planets could be overestimated by as much as 50\%. We also show that
whenever high-resolution imaging reveals a transit host star to be a
binary, the planet is usually more likely to orbit the primary star
than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

A lingering concern about these studies is that in most cases,
investigators have assumed that all of the targets in the survey
are single stars~\citep[\textit{e.g.},][]{
  howard_planet_2012,fressin_false_2013,
  dressing_occurrence_2015,burke_terrestrial_2015}. 
In reality, many of the sources that are monitored in a transit survey are
unresolved multiple-star systems, mainly binaries.  Unrecognized
binaries cause numerous systematic \deleted{errors}\added{uncertainties} in 
the planetary occurrence
rates.  For example, when there is a transiting planet around a star
in a binary, the additive constant light from the second star reduces
the fractional loss of light due to the planet.  This makes transit
signals harder to detect and lowers the number of detections.  On the
other hand, a binary system presents two opportunities to detect
transiting planets, which could increase the overall number of
detections.

@@

There are many potential pitfalls in this calculation.  Some genuine
transit signals are missed even if they formally exceed the
signal-to-noise threshold, because of the probabilistic nature of
transit detection.  Planets can be misclassified due to statistical
and systematic \deleted{errors}\added{uncertainties} in the catalogued 
properties of the stars.  Some
transit-like signals are spurious, arising from noise fluctuations or
failures of ``detrending'' the astrophysical or instrumental
variations in the photometric signal.  Poor angular resolution leads
to blends between eclipsing binary stars and other stars along nearly
the same line of sight, producing signals that mimic those of
transiting planets.

Here, though, we will focus exclusively on problems that arise from
the fact that many stars exist in unresolved binary systems.  We will
also focus on the errors in planet occurrence as a function of radius,
rather than orbital period. This is because when more than one transit
is detected (as is usually required by the surveyors), the orbital
periods can be measured without ambiguity regardless of whether the
host star is single or one member of a binary.  Even with this narrow
focus, there are numerous sources of error.  All three of the
quantities in Equation~\ref{eq:occ_rate_simple} are biased:
\begin{enumerate}
%    
    \item The number of detected planets, $N_{\det}$, is actually the
      number of detected planets that {\it appear} to have size $r$,
      orbiting stars that {\it appear} to have mass $M$.  Whenever the
      planet-hosting star is part of a binary,
%    
    \begin{itemize}
        \item the planet's size could be misclassified because of the
          reduction in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified
          because its light is combined with a second star of a
          different spectral type.
%        
    \end{itemize}
%    
    \item The number of searchable stars, $N_\star$, is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of
          the secondary stars that were inadvertently searched for
          transiting planets;
%        
        \item toward higher values, because some of the stars that
          appeared to be searchable are in fact binaries for which the
          amplitude of the photometric signal would have been reduced
          to an undetectable level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$ is biased because the
      planet-hosting star could be misclassified.  At fixed orbital
      period, the transit probability is proportional to
      $\rho^{-1/3}$, where $\rho$ is the stellar mean
      density~\citep{Winn2010}.  Therefore, any 
      \deleted{errors}\added{uncertainties} in determining
      the host star's mean density lead to 
      \deleted{errors}\added{uncertainties} in the correction
      for the transit probability.
%    
\end{enumerate}

@@

Since binarity produces \deleted{the largest}\added{large} errors in inferred
planetary radii when the two stellar components are similar, we
begin by considering a universe in which
all stars are identical, with mass $M$, radius $R$, and luminosity
$L$.  Single stars are uniformly distributed in space with a number
density of $n\s$ stars per cubic parsec, and binary systems are
uniformly distributed with number density $n\b$.  In this scenario,
stars are never misclassified because the combined light of a binary
has the same color and spectrum as a single star.  We further assume
that all planets have the same radius, $r$, and occur around single
stars and members of binaries at the same rate, $\Lambda_r$.

@@

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=0.6\textwidth]{f1.pdf}
    \end{center}
    \caption{ {\it Top.}---Volumes within which sources appear to be
    searchable for planets of radius $r$.  Single stars are searchable
    out to a distance $d_0$, at which point they become too faint to
    allow the detection of transits.  Binaries are brighter and
    therefore appear to be searchable out to a larger distance.  {\it
    Bottom.}---Volumes within which stars appear to be searchable for
    planets of radius $r/\sqrt{2}$.  }
    \label{fig:model_1_volumes}
\end{figure}

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=.6\textwidth]{f2.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{ Apparent occurrence rate $\Lambda\a$, and true
    occurrence rate $\Lambda$ in a universe with only one type of
    star, one type of planet (with radius $r$), and $n\b/n\s=0.05$,
    for $n\b$ the number density of binary systems, and $n\s$ the
    number density of single stars.  The occurrence rate of planets
    with radius $r$ is underestimated, and the occurrence rate of
    planets with radius $r/\sqrt{2}$ is overestimated.}
    \label{fig:occ_rate_model_1}
\end{figure}

@@

\input{case_study_toc.tex}

We can now gauge the size of the systematic 
\deleted{errors}\added{uncertainties} associated
with unresolved binaries, for \deleted{cases of interest.}\added{the cases of 
interest listed in Table~\ref{tab:case_study_toc}.}
Throughout this
section we assume $R\propto M$, and $L \propto M^{\alpha}$ where
$\alpha = 3.5$, as is roughly the case for main-sequence stars.
With these assumptions, Equations~\ref{eq:dil_1} and~\ref{eq:dil_2}
become
\begin{equation}
  \mathcal{D}_1
  =  (1+q^\alpha)^{1/2}~~{\rm and}~~\mathcal{D}_2
  = q(1+q^{-\alpha})^{1/2}.
\end{equation}

@@

\begin{figure}[!tb]
    \centering
    \includegraphics[width=0.6\textwidth]{f3.pdf}
    \caption{ The mass ratio distribution of binaries in a
    magnitude-limited sample, assuming the underlying volume-limited
    distribution is uniform and the luminosity-mass relation is
    $L\propto M^{3.5}$. For transit signals of a given amplitude, the
    sample of searchable sources is magnitude-limited, causing high
    mass ratios to be over-represented. }
    \label{fig:q_distribn_mag_limited}
\end{figure}

@@

Again assuming $\Gamma_0=\Gamma_1=\Gamma_2$, the apparent occurrence
rate density is 
\begin{align}
    %see annoying_q_factor_dec22.pdf for derivation
    \notag
    %\frac{\left.\Gamma\a(r\a)\right|_{r\a= r}}{\Gamma_0(r)} 
    \frac{\Gamma\a(r\a)}{\Gamma_0(r\a)} 
    &=
    \frac{1}{1+\mu}
    \left[
      1 + \frac{1}{\mathcal{N}_q} \frac{n\b}{n\s}
    \left(
      \int_0^1 {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\gamma+4}{2}} +
      \right.
      \right. \\
      %isn't latex great?
      &\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
      \left.\left.
      \int_0^1 {\rm d}q\,q^{\beta+\gamma+\frac{5}{3}} 
      (1+q^\alpha)^{\frac{3}{2}}
      (1+q^{-\alpha})^{\frac{\gamma+1}{2}}
    \right)
    \right],
    \label{eq:powerlaw_vary_binary}
\end{align}
Adopting the realistic numerical values $n\b/n\s=0.79$, $\alpha =
3.5$, $\beta=0$, and $\gamma=-2.92$, the summed integrals in
Equation~\ref{eq:powerlaw_vary_binary} give $(\ldots)\approx 1.503$,
and $\Gamma\a/\Gamma_0 = 1.048$.  The apparent occurrence rate density
is 4.8\% higher than the true value.
\added{Exploring different values for the power-law index,
we find that $\Gamma\a/\Gamma_0$ lies between $0.99$ and 
$1.10$, for $\gamma=-3.5$ and $\gamma=-2.5$, respectively.}
We conclude that for planets around
Sun-like stars with orbits within 0.25\,AU and radii in the range
2--17\,$R_\oplus$, the systematic \deleted{errors}\added{uncertainties} 
associated with binarity are
on the order of a few percent.

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{f4a.pdf}{f4b.pdf}
    \caption{ {\it Left.}---Apparent occurrence rate density
      ($\Gamma\a$), compared to the true occurrence rate density
      ($\Gamma_0$) for single stars and primary stars in binaries.
      Three different cases are plotted, with different choices for
      the number of planets per secondary star relative to
      that for single stars ($Z_2/Z_0$).
      {\it Right.}---Same, integrated over $0.5\,R_\oplus$
      bins to give apparent and true occurrence rates.
      In both cases the true planet
      radius distribution is specified by
      Equation~\ref{eq:model3_radius_distribution}.
    }
    \label{fig:occ_rate_model_3_log}
\end{figure}

For planets smaller than 2\,$R_\oplus$, the true occurrence rate
density is more uncertain because such planets are more difficult to
detect.  To investigate the systematic \deleted{errors}\added{uncertainties} 
associated with binarity
on the inferred occurrence rate of Earth-sized planets, we need to
make plausible assumptions about the true occurrence rate density.
Obviously if the rate density continues to vary as $r^{-3}$ down to
much smaller planet sizes, the results of the preceding section will
hold.  But there is no particular reason to think this will be the
case and indeed, some investigators have concluded that the rate
density begins to level off to a constant value as $r$ decreases below
$2R_\oplus$~\citep{petigura_plateau_2013}.  To investigate the
implications we consider a broken power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2R_\oplus \\
        {\rm constant} & \text{for } r\leq2R_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
In this case the integrals that appear in
Equation~\ref{eq:general_Gamma_a} are tedious to work out
analytically, leading us to evaluate them numerically.\footnote{We
refer the interested reader to our online code for performing this
computation: \url{github.com/lgbouma/binary_biases}, commit
\texttt{fd41b4b}.} Figure~\ref{fig:occ_rate_model_3_log} shows the
results, again for the case $n\b/n\s=0.79$, $\alpha = 3.5$, $\beta=0$,
and $\gamma=-2.92$.

@@

\subsection{Which star has the planet?}
\label{subsec:whichstar}
\begin{figure}[!t]
    \centering
    \includegraphics[width=\textwidth]{f5.pdf}
    \caption{ Fraction of detected planets that orbit single stars
      (solid lines), primary stars of binaries (dashed lines), and
      secondary stars of binaries (dotted lines), as a function of the
      apparent radius of the planet.  We assume that single stars and
      primaries have the same number of planets per star.  Three
      different cases are plotted, differing in the relative
      occurrence of planets around secondary stars ($Z_2/Z_0$).  The
      true radius distribution is given by
      Equation~\ref{eq:model3_radius_distribution}.}
    \label{fig:frac_model_3}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{f6a.pdf}{f6b.pdf}
    \caption{ {\it Left.}---True and apparent occurrence rate densities,
      and {\it right.}---occurrence rates, as a function of planet
      radius, using a model for the radius distribution that exhibits
      a gap between 1.5 and 2\,$R_\oplus$
      (Equation~\ref{eq:model4_radius_distribution}). 
      \added{The peak in apparent occurrence rate density between 1.3 and 
      1.5$\,R_\oplus$ (solid dark blue line, relative to dotted light blue 
      line) is the contribution from secondaries with true radii greater
      than 2\,$R_\oplus$.}
    }
    \label{fig:model_4}
\end{figure}
